---
title: "HW9"
author: "Kylie Knudten"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## HW9

### Group members: Kylie Knudten, Alison Wyatt

Github link: https://github.com/aawyatt/HW9

## Part 1: Data tidying

```{r}
# import libraries
library(dplyr)
library(tidyverse)
# read in data
av <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/avengers/avengers.csv",stringsAsFactors = FALSE)

names(av)
av0 = av %>% filter(Name.Alias != "")
av1 = av0 %>% gather(key = time, value = death, c(11, 13, 15, 17, 19)) %>% 
  select(Name.Alias, time, death) %>%
  mutate(time = parse_number(time))
av2 = av0 %>% gather(key = time, value = return, c(12, 14, 16, 18, 20)) %>% 
  select(Name.Alias, time, return) %>%
  mutate(time = parse_number(time))
av.neat = left_join(av1, av2, by = c("Name.Alias", "time"))
av0 = av0 %>% filter(Name.Alias != "Vance Astrovik")
av1 = av0 %>% gather(key = time, value = death, c(11, 13, 15, 17, 19)) %>% 
  select(Name.Alias, time, death) %>%
  mutate(time = parse_number(time))
av2 = av0 %>% gather(key = time, value = return, c(12, 14, 16, 18, 20)) %>% 
  select(Name.Alias, time, return) %>%
  mutate(time = parse_number(time))
av.neat = left_join(av1, av2, by = c("Name.Alias", "time"))
av.neat %>% group_by(time) %>% summarise(total.num = n(),
                                         real.death = sum(death == "YES"), 
                                         fake.death = sum(death == "NO"), 
                                         missing = sum(death == ""))
av.neat %>% filter(time == 1, death == "YES") %>% 
  summarise(total.num = n(),
            real.return = sum(return == "YES"), 
            fake.return = sum(return == "NO"), 
            missing = sum(return == ""))
View(av.neat)
```

## Part 2: Validation

### Kylie's statement: "There's a 2-in-3 chance that a member of the Avengers returned from their first stint in the afterlife"

```{r}
av.neat %>% filter(time==1, death=="YES") %>% summarise(total.num = n(), return1death = sum(return == "YES"), chance = return1death/total.num)
```

This statement is true, as the total percent of Avengers who returned after their first death is 0.698, which is just slightly higher than a 2-in-3 chance which is around 0.667.


### Alison's statement: "Of the nine Avengers we see on screen — Iron Man, Hulk, Captain America, Thor, Hawkeye, Black Widow, Scarlet Witch, Quicksilver and The Vision — every single one of them has died at least once in the course of their time Avenging in the comics...all those top-tier Avengers get better after their first death"

```{r}
av.neat %>% filter(Name.Alias %in% c("Anthony Edward \"Tony\" Stark", "Robert Bruce Banner", "Steven Rogers", "Thor Odinson", "Clinton Francis Barton", "Natalia Alianovna Romanova", "Wanda Maximoff", "Pietro Maximoff", "Victor Shade (alias)"), time == 1) %>% select(Name.Alias, time, death, return)
```
All of the on screen Avengers did in fact die once - and also returned after their first death.
